\section*{Acknowledgements}
Everyone who made improvements to this document...

\pagebreak
\bibliographystyle{plain}

\begin{thebibliography}{25}

\bibitem{Ashida}
Ashida K. and Mishiue M., 1973. \emph{Studies on bed load transport
rate in alluvial streams}, Trans JSCE, vol.4.

\bibitem{Bailard}
Bailard J., 1981. An energetics total load transport model
for a plane sloping beach. \emph{Journal of Geophysical Research}, 86, C11,
10938-10954.

\bibitem{Begnudelli}
Begnudelli L., Valiani A. and Sanders B.F., 2010. A balanced treatment of secondary currents, turbulence and dispersion in a depth-integrated hydrodynamic and bed deformation model for channel bends. \emph{Advances in Water Resources}, 33(1).

\bibitem{Belleudy}
Belleudy P., 2000. Numerical simulation of sediment mixture
deposition, part 1: analysis of a flume experiment, \emph{Journal of Hydraulic
Research}, Vol.38, N$^\circ$6.

\bibitem{Bijker}
Bijker E.W., 1968. Mechanics of sediment transport by the
combination of waves and current. In \emph{Design and Reliability of Coastal
Structures}, 23$^{rd}$ Int. Conf. on Coastal Engineering, 147-173.

\bibitem{Camenen}
Camenen B., 2002. \emph{Mod\'{e}lisation num\'{e}rique du transport s%
\'{e}dimentaire sur une plage sableuse}. Ph.D thesis, Universit\'{e}
Joseph Fourier, Grenoble 1.

\bibitem{CelikRodi}
Celik I., Rodi W., 1988. Modelling Suspended Sediment Transport
in non- equilibrium situations, \emph{Journal of Hydraulic Eng.}, Vol. 114, N$^\circ
$10.

\bibitem{Chollet}
Chollet J.P. and Cunge J.A., 1980. New
interpretation of some headlooss - flow velocity relationship for deformable
bed, \emph{J. Hydr. Eng.}, 17 (1).

\bibitem{DaviesVillaret}
Davies A.G. and Villaret C., 2004. \emph{Modelling the effect of
wave-induced ripples on littoral sand transport}. EDF-LNHE Report
HP-75/03/029/A.

\bibitem{Dibajnia}
Dibajnia M., Watanabe A., 1992. \emph{Sheet flow under non-linear
waves and currents}. Proc. of the Intl. Conf. on Coast Eng., 2015-2029.

\bibitem{Egiazaroff}
Egiazaroff I.V., 1965: Calculation of non-uniform sediment
concentrations, \emph{J. of Hydr. Div. ASCE}, vol. 91, N$^\circ$4, pp. 225-248.

\bibitem{Einstein}
Einstein H.A. (1950). \emph{The bed load function for sediment
transportation in open channel flow}. US Dep. Of Agriculture, Techn. Bull.

\bibitem{Engelund}
Engelund F. and Hansen E., 1967. \emph{A monograph on sediment transport
in alluvial streams}. Techn. Univers. Of Denmark,
Copenhagen, Denmark. 

\bibitem{Gonzales}
Gonzales de Linares M., 2002. \emph{Graded
sediment in Sisyphe}. Rapport EDF-LNHE, HP-75/2002/68.

\bibitem{Hirano}
Hirano, M., 1970. \emph{On phenomena of river-bed lowering and
armouring below reservoirs}. In Proc. 14th Hydraul. Lecture Meeting, 13-14
Feb., Civ. Eng. Ass. Hydr. Committee, Hatsumei Kaikan.

\bibitem{JMH2}
Hervouet J.-M., \emph{Hydrodynamics of free surface flows: modelling with finite element methods}, John Wiley \& Sons, 2007.

\bibitem{Hervouet11}
Hervouet J.M., Razafindrakoto E., Villaret C., 2011. \emph{Dealing with dry zones
in free surface flows, a new class of advection scheme}. IAHR conference, July 2011.

\bibitem{Hunziker}
Hunziker R.P., 1995. \emph{Fraktionsweiser Geschuebetransport},
Ph.D. thesis, Mitteeilungen Nr 138 deer Versuchanstalt fur Wasserbau,
Hydrologie und Glaziologie, ETH Zurich, Switzerland.

\bibitem{Huybrechts}
Huybrechts N., Villaret C., Hervouet JM, 2010. \emph{Comparison between 2D and 3D
modelling of sediment transport: application to the dune evolution}. Proceedings of the RiverFlow conference.

\bibitem{Huynh}
Huynh-Thanh and Temperville A., 1991. \emph{A numerical model of the
rough turbulent boundary layer in combined wave and current interaction}. In
Sand Transport in Rivers Estuaries and the Sea, eds RL Soulsby and R.
Bettess, pp. 93-100, Balkema, Rotterdam.

\bibitem{Karim}
Karim M.F. and Kennedy J.F., 1982. \emph{A computer based flow and
sediment routing}. IIH Report N$^\circ$250 Modelling for streams and its
application to the Missouri River, University of Iowa, Iowa City, IA.

\bibitem{Koch}
Koch F.G. and Flokstra C., 1981. \emph{Bed level computations for
curved alluvial channels}, XIXth Congress of the International Association
for Hydraulic Research, New Delhi India.

\bibitem{Meyer-Peter}
Meyer-Peter E. and Muller R., 1948. \emph{Formulae for bed-load
transport}. Int. IARH Congress, Stockholm, Sweden. 

\bibitem{Ribberink}
Ribberink J., 1987. \emph{Mathematical modelling of one-dimensional morphological
changes in rivers with non-uniform sediment}. PhD thesis, Delft University.

\bibitem{Roelvink06} Roelvink J.A., \emph{Coastal morphodynamic evolution techniques}, Coastal Engineering, 2006, 53, 277-287.

\bibitem{Soulsby}
Soulsby R., 1997. \emph{Dynamics of marine sands}. Thomas Thelford Edition.

\bibitem{Swart}
Swart D.H., 1976. \emph{Offshore sediment transport and
equilibrium beach profiles}. Delft Hydraulics Publication 131, Delft University,
The Netherlands.

\bibitem{Tanaka}
Tanaka, H. and Dang, V.T., 1996. Geometry of sand ripples due
to combined wave-current flows. \emph{J. of Waterway, Port, Coastal and Ocean
Engineering}, ASCE, 122 (6), 298-300.

\bibitem{vanRijn84}
van Rijn L.C., 1984. Sediment transport - Part I : bed load -
Part II: suspended load, \emph{J. of Hydraulic Division}, Proc. ASCE, 110, HY10,
1431-56, HY11, 1613-41.\\

\bibitem{vanRijn87}
van Rijn L.C., 1987. \emph{Mathematical modelling of morphological processes in
the case of suspended sediment transport}, Doctoral Thesis, Faculty of civil
engineering, Delft University of technology).

\bibitem{vanRijn93}
van Rijn L.C., 1993. \emph{Principles of Sediment Transport in
Rivers, Estuaries and Coastal Seas}. Aqua Publications.

\bibitem{Villaret01a}
Villaret C., 2001. \emph{Prise en compte des fonds non-\'{e}rodables
dans le logiciel Sisyphe}, Rapport EDF-LNH, HP-75/02/045/B.

\bibitem{Villaret01b}
Villaret C., 2001. \emph{Mod\'{e}lisation du transport solide par la
formule de Bijker-Etude de sensibilit\'{e} et tests validation}. Rapport EDF
R\&D LNHE HP-75/2001/66/A.

\bibitem{Villaret04}
Villaret C. and Davies A.G., 2004. \emph{Numerical modeling of
littoral sand transport}. ICCE Conference, Portugal, September.

\bibitem{Wiberg}
Wiberg P.L. and Harris C.K., 1994. Ripple geometry in
wave-dominated environments. \emph{Journal of Geophysical Research} 99 (C1),
775-789.

\bibitem{Waeles}
Waeles B., 2005. \emph{Detachment and transport of slay sand gravel
mixtures by channel flows}. Ph.D. thesis, University of Caen.

\bibitem{Yalin}
Yalin, MS, 1977. \emph{Mechanics of sediment transport}. Pergamon
Press, Oxford.

\bibitem{YalinFerrera}
Yalin M.S. and Ferrera da Silva A.M.F., 2001. \emph{Fluvial
Processes}. IAHR monograph Delft Hydraulics.

\bibitem{Wu}
Wu W., 2008. \emph{Computational River Dynamics}. Taylor and Francis Group, London, UK.

\bibitem{Zyserman}
Zyserman J.A. and Fredsoe J. 1994. Data analysis of bed concentration of
suspended sediment, \emph{Journal of Hydraulic Engineering}, ASCE, Vol. 120, N$%
^\circ$9, pp 1021-1042.

\end{thebibliography}




