Pham Van Bang D., Lefran\c{c}ois E., Sergent P., Bertrand F., 2008. MRI
experimental and finite elements modelling of the
sedimentation-consolidation of mud, La Houille Blanche, 168, n$^\circ$%
3-2008, 39-44.\newline
Sanchez, M., 1992, Mod\'{e}lisation dans un estuaire \`{a} mare: R\^{o}le du
bouchon vaseux dans la tenue des sols sous marins. PhD thesis. University of
Nantes (232 pages).\newline
Bartholomeeusen, G., Sills, G.C., Znidarcic, D., van Kesteren, W.,
Merckelbach, L.M., Pyke, R., Carrier, W.D., Lin, H., Penumadu, D.,
Winterwerp, H., Masala, S. and Chan, D., 2002, Sidere: numerical prediction
of large strain consolidation, G\'{e}otechnique 52, No. 9, 639-648.\newline
Been, K. and Sills, G.C., 1981, Self weight consolidation of Soft Soils: an
Experimental and Theoretical Study. Geotechnique, Volume 31, No.4, pp.
519-535.\newline
Camenen B. \& Pham Van Bang D., 2011. Modelling the settling of suspended
sediments for concentrations close to the gelling concentration, Cont. Shelf
Res., 31, S106-S116.\newline
Hervouet, J.M. 2007. Hydrodynamics of free surface flow, modelling with
finite elements system. Wiley. 341p.\newline
Gibson, R.E., Englund, G. L. and Hussey, M. J. L., 1967, The theory of one
dimensional consolidation of saturated clay, I. Finite Non-Linear
Consolidation of Thin Homogeneous Layers. Geotechnique, pp. 261-273\newline
Gibson, R.E., Schiffman, R.L. and Cargill, K.W., 1981, The theory of One
dimensional Consolidation of Saturated Clays, II. Finite Nonlinear
Consolidation of Thick Homogeneous Layers, Canadian Geotechnical Journal,
Vol. 18, pp. 280-293\newline
Hervouet J.M. (2007): Hydrodynamics of Free Surface Flows, modelling with
the finite- element method, J. Wiley \& Sons Ltd, West Sussex, England, 340pp%
\newline
Lenormant, C., Lepeintre, F., Teisson, C., Malcherek, A., Markofsky, M., and
Zielke, W., 1993, Three dimensional modelling of estuarine processes. In
MAST Days and Euromar Market, Project Reports Volume 1.\newline
Leveque, Randal J., 2002, Finite-Volime Methods for Hyperbolics Problems,
Cambridge Texts in Applied Mathematics, ISBN 0-511-04219-1 eBook.\newline
Migniot C., 1968. A study of the physical properties of different very fine
sediments and their behaviour under hydrodynamic action, La Houille Blanche
7, 591-620.\newline
Partheniades, E., 1962, A study of erosion and deposition of cohesive soils
in salt water. Ph.D thesis, University of California, Berkeley, 182 p.

Pham Van Bang D., Ovarlez G., Tocquer L., 2007. Density and structural
effects on the rheological characteristics of mud, La Houille Blanche, 2,
85-93.

Tassi, P., Villaret C., Huybrechts,N. Hervouet, JM. (2011): Numerical
modelling of 2D and 3D suspended sediment transport in turbulent flows,
Proceedings of the RCEM 2011 Conference in Beijing.\newline
Pham Van Bang D., Lefran\c{c}ois E., Sergent P., Bertrand F., 2008. MRI
experimental and finite elements modelling of the
sedimentation-consolidation of mud, La Houille Blanche, 168, n$^\circ$%
3-2008, 39-44.\newline
Sanchez, M., 1992, Mod\'{e}lisation dans un estuaire \`{a} mare: R\^{o}le du
bouchon vaseux dans la tenue des sols sous marins. PhD thesis. University of
Nantes (232 pages).\newline
Terzaghi, K. 1923. Die Berechnung des Durchlassigkeitsziffer des Tones aus
des Verlauf des hydrodynamischen Spannungerscheinungen, Sitz. Akad. Wissen.
Wien, Math. Naturwiss. Kl., Abt IIa, 132, 125-138.\newline
Thi\'{e}bot J., 2008, Mod\'{e}lisation num\'{e}rique des processus
gouvernant la formation et la d\'{e}gradation des massif vaseux (PhD thesis
ENGREF- U.Caen), 130 p. \newline
Thiebot, J., Guillou, S., Brun-Cottan, J-C., 2011, An optimisation method
for determining permeability and effective stress relationships of
consolidating cohesive sediment deposits, Continental Shelf Research 31
(2011) S117-S123.\newline
Toorman E. A., 1996, Sedimentation and self-weight consolidation: general
unifying theory, G\'{e}otechnique 46, N$^\circ$ 1, pp. 101-113\newline
Toorman, E.A., 1999. Sedimentation and self-weight consolidation:
constitutive equations and numerical modelling. G\'{e}otechnique,
49(6):709-726.

Van L.A., 2012: Mod\'{e}lisation du transport des sediments mixtes
sable-vase et application \`{a} la morphodynamique de l'estuaire de la
Gironde, PhD Thesis, Universit\'{e} Paris-Est.\newline
Van L.A., Villaret C., Pham van Bang D., Sch\"{u}ttrumpf H., 2012, Erosion
and deposition of the Gironde mud, International Conference on Scour and
Erosion, ICSE-6 -- Paris- August 27-31, 2012.\newline
Van L.A., Pham Van Bang D., 2012, Hindered settling of sand/mud flocs
mixtures: from model formulation to numerical validation, Advances in Water
Resources (accepted).\newline
Van Leussen, 1994: Estuarine macroflocs and their role in fine grained
transport, PhD Thesis, Utrecht University(Nederlands).

Villaret, C., Tassi, P. 2012{\nobreakspace}: SISYPHE User manual, release
6.2,. Rapport EDF-LNHE H-P74-2012-02004-EN .\newline
Villaret C., Walther R. ,2008: Numerical modeling of the Gironde estuary.
Physics of Estuaries and Coastal Sediments, Liverpool, August 2008.\newline
Villaret C., Van L.A., Huybrechts N., Pham Van Bang D., Boucher O. 2010.
Consolidation effects on morphodynamics modelling: application to the
Gironde estuary, La Houille Blanche, N$^\circ$ 6-2010, 15-24.\newline
Winterwerp, J.C. and van Kesteren, W.G.M., 2004, Introduction to the physics
of cohesive sediment in the marine environment, Development in Sedimentology
56, Elsevier (466 pages).

Waeles B., 2005: Detachment and transport of slay sand gravel mixtures by
channel flows, Ph. D. Thesis University of Caen.

