\chapter{The inputs / outputs}

\section{Preliminary comments}

In a computation, the \telemac{3D} code uses a number of input and output files,
some of which are optional. Most of these files are similar or identical to
their counterparts in \telemac{2D}.

The input files are:

\begin{itemize}
\item the steering file (mandatory), which contains the "configuration" of the
simulation,

\item the geometry file (mandatory), which contains the information regarding
the mesh,

\item the boundary conditions file (mandatory), which contains the description
of the type of each boundary,

\item the Fortran file, which contains the specific subroutines of the
simulation (modified TELEMAC subroutine or specifically created),

\item the bottom topography file, which contains the description of the bottom
topography. Generally, the topographic data are already contained in the
geometry file and the bottom topography file is generally not used,

\item the liquid boundaries file, which contains the information on the
imposed values at liquid boundaries,

\item the previous computation file, which provides the initial state of the
calculation in the case of a continuation calculation,

\item the reference file, which contains the calculation of "reference" used
for the validation process,

\item the stage-discharge curves file, which contains the information on the
imposed values at liquid boundaries in case of height / flow rate law,

\item the sources file, which contains the information regarding the sources.
\end{itemize}

The output files are:

\begin{itemize}
\item the 3D result file, which contains the graphical results associated to
the 3D mesh,

\item the 2D result file, which contains the graphical results associated to
the 2D mesh,

\item the output listing, which is simulation report. In case of difficulty in
performing a calculation, the user can request the printing of additional
information by activating the logical keyword \telkey{DEBUGGER}.
\telkey{DEBUGGER = 1} provides the calling sequences of subroutines from the
main program telemac3d.f. This technique is useful in case of critical
computation crash to identify the responsible subroutine,

\item the file for Scope, which is an additional test file available for the
user,

\item the restart file, allowing to perform a continuation computation without
information loss.
\end{itemize}

In addition, the user may have to manage additional files:

\begin{itemize}
\item the binary data files 1 and 2, in input,

\item the formatted data files 1 and 2, in input,

\item the binary results file, in output,

\item the formatted results file, in output.
\end{itemize}


\subsection{Binary files format}
\label{sec:binfile}
the binary files used within the TELEMAC system can have various formats. The
most common format is the SERAFIN format (also known as SELAFIN, for
misidentified historical reasons) which is the standard internal TELEMAC system
format (described in Appendix N${}^\circ$ 3). This SERAFIN format can be
configured so as to store the real-typed values in single or double precision.
The other available format is the MED format which is used by the Salomé
platform jointly developed by EDF and the French atomic energy commission
(CEA). In the current TELEMAC release, this format is restricted to EDF
internal use.

Depending on the specified format, the binary files may be treated by different
software. However, in the current release of TELEMAC, only the SERAFIN single
precision format can be read by the post-processing tools like RUBENS,
FUDAA-PREPRO or BLUE KENUE.

The selection of the format of a file is done by the corresponding keyword.
Thereby, the keyword \telkey{GEOMETRY FILE FORMAT} specifies the format of the
geometry file. Those keywords may take 3 different values (as an 8 character
string): the value 'SERAFIN' corresponds to the standard single precision
SERAFIN format, which is the default and recommended value (do not forget the
space in last position). The 'SERAFIND' value corresponds to the double
precision SERAFIN format which allows to increase the precision of the results,
especially in case of a restart or reference file. Finally, the 'MED' value
corresponds to the hdf5 MED format (internal EDF use).

The keywords involved are:

\begin{itemize}
\item \telkey{2D RESULT FILE FORMAT},

\item \telkey{3D RESULT FILE FORMAT},

\item \telkey{BINARY DATA FILE 1 FORMAT},

\item \telkey{FILE FOR 2D CONTINUATION FORMAT},

\item \telkey{GEOMETRY FILE FORMATGEOMETRY FILE FORMAT},

\item \telkey{PREVIOUS COMPUTATION FILE FORMAT},

\item \telkey{REFERENCE FILE FORMAT},

\item \telkey{RESTART FILE FORMAT}.
\end{itemize}

\section{The steering file}

the steering file contains all the data about the selection of computational
options (physical, numerical, etc.). It is an ASCII file which can be generated
either through the FUDAA-PREPRO software or directly using a text
editor. In a way, it serves as the computation dashboard. It includes a set of
keywords to which values are assigned. If a keyword does not occur in that
file, then \telemac{3D} will assign it the default value as defined in the
dictionary file. If such a default value is not defined in the dictionary, then
the computation will be interrupted with an error message. For instance, the
command \telkey{TIME STEP = 10.0} specifies that the computation time step
value is 10 seconds.

\telemac{3D} reads the steering file at the beginning of the computation.

Both dictionary file and steering file are read using the DAMOCLES library
which is included in the TELEMAC chain. It is therefore necessary, when
generating the steering file, to observe the DAMOCLES syntax rules (what is
performed automatically if the file is generated using FUDAA-PREPRO).

The writing rules are as follows:

\begin{itemize}
\item the keywords can be of the Integer, Real, Logical or Character type,

\item the sequence order in the steering file is of no importance,

\item several keywords can be on the same line,

\item each line cannot exceed 72 characters. However, one can start a new line
as many times as one wishes provided that the keyword name is not astride two
lines,

\item For the array-types keywords, the character separating successive values
is the semicolon. For example:

\begin{lstlisting}[language=TelemacCas]
PRESCRIBED FLOWRATES =   10.0;20.0
\end{lstlisting}

\item The symbols ":" and "=" can both be used in order to separate a keyword
name and its value. They can be either preceded or followed by any number of
blanks. The value itself can occur on the following line. For example:

\begin{lstlisting}[language=TelemacCas]
TIME STEP =   10.
\end{lstlisting}

or

\begin{lstlisting}[language=TelemacCas]
TIME STEP : 10.
\end{lstlisting}

Or else

\begin{lstlisting}[language=TelemacCas]
TIME STEP =
10
\end{lstlisting}

\item The characters occurring between two "/" on one line are taken as
comments. Likewise, those characters occurring between a "/" and a line ending
are considered as comments. For example:

\begin{lstlisting}[language=TelemacCas]
/ k-epsilon model
VERTICAL TURBULENCE MODEL = 3
\end{lstlisting}

\item A line beginning with a "/" in a first column is wholly considered as a
comment, even though there is another "/" on the line. For example:

\begin{lstlisting}[language=TelemacCas]
/ The geometry file is ./mesh/geo
\end{lstlisting}

\item Writing the integers: do not exceed the maximum size allowed by the
machine (in a machine with a 32 bit architecture, the extreme values range from
-2~147~483~647 to +2~147~483~648. Do not insert any blank between the sign
(optional for the sign +) and the number. A point after the end of the number
is tolerated,

\item Writing the reals: the point and the comma are accepted as a decimal
separator, as well as the Fortran formats E and D. ( 1.E-3  0.001  0,001  1.D-3
represent the same value),

\item Writing the logical values: the values 1, OUI,  YES,  .TRUE.,  TRUE,  or
VRAI on the one hand, and 0, NON,  NO,  .FALSE.,  FALSE, or FAUX on the other
hand, are accepted,

\item Writing the character strings: the strings including blanks or reserved
symbols ("/",":", "=", "\&") should be inserted between quotes ('). The value
of a character keyword may contain up to 144 characters. As in Fortran, the
quotes included in a string should be doubled. A string may not begin or end
with a blank. For example:

\begin{lstlisting}[language=TelemacCas]
TITLE = 'COASTAL ENVIRONMENT STUDY'
\end{lstlisting}

\end{itemize}

In addition to the keywords, a number of pseudo instructions or metacommands
which are interpreted during the sequential reading of the steering file can
also be used:

\begin{itemize}
\item the \telkey{\&FIN} command specifies the file end (even though the
file is not over). Thus, some keywords can be disabled simply by placing them
behind that command so that they can easily be reactivated later on. However,
the computation keeps running,

\item the \telkey{\&ETA} command prints the list of keywords and the
values which are assigned to them when DAMOCLES meets that command. That
display will take place at the head of the output listing,

\item the \telkey{\&LIS} command prints the list of keywords. That
display will take place at the head of the output listing,

\item the \telkey{\&IND} command prints the detailed list of keywords.
That display will take place at the head of the output listing,

\item the \telkey{\&STO} command causes the program to be halted, whereas
the computation does not keep running.
\end{itemize}

\section{The geometry file}

it is the same file as that used by \telemac{2D}. It is a SERAFIN-formatted
binary file which can then be read by RUBENS or FUDAA-PREPRO and is generated
either by MATISSE, BLUE KENUE or JANET or by the \stbtel module (from the
file(s) originating from of mesh generator). The SERAFIN format structure is
described in Appendix \ref{sec:srffmt}.

That file contains all the data about the two-dimensional mesh (see in
subsection 2.3.2). It includes the number of points in the mesh (NPOIN2NPOIN2
variable), the number of elements (NELEM2 variable), the number of vertices per
element (NDPNDP variable), the X and Y arrays containing the co-ordinates of
all the points and, lastly, the IKLE array containing the connectivity table.

That file may also contain bathymetric and bottom friction data for each point
in the mesh.

NOTE: \telemac{3D} retrieves the geometry data at the beginning of the 2D result
file. Any computational 2D result file can therefore be used as a geometry file
when a further simulation on the same mesh is desirable.

That file name is provided using the keyword: \telkey{GEOMETRY FILE} and its
format is specified by the keyword: \telkey{GEOMETRY FILE FORMAT} (default
value: 'SERAFIN '), see subsection \ref{sec:binfile}.


\section{The boundary conditions file}

it is the same file as that used by \telemac{2D}. It is a formatted file which is
generated automatically by MATISSE or \stbtel and can be amended
using FUDAA-PREPRO or a text editor. Each line in that file is
dedicated to a point at the 2D mesh boundary. The edge point numbering is that
of the file lines; it first describes the domain outline in the counter
clockwise direction from the bottom left-hand point (point where the sum of
horizontal co-ordinates of which is minimum), then the islands in the clockwise
direction.

For a thorough description of that file, refer to the specific subsection
\ref{sec:bndfile}.

That name of file is provided using the keyword: \telkey{BOUNDARY CONDITIONS
FILE}.


\section{The fortran file}

the Fortran file may include a number of subroutines (so-called "user"
subroutines) available under the \telemac{3D} tree structure which the user can
modify as well as those subroutines which have been specifically developed for
the computation.

The user subroutines from the various libraries used by \telemac{3D} are listed
in Appendix \ref{sec:usrsub}. Since \telemac{3D} is available in Open Source, every
subroutine can be freely used. Every user subroutine copied into the user
Fortran file is automatically substituted for the same named subroutine
occurring in the \telemac{3D} compiled libraries.

Upon the creation and every amendment of the Fortran file, a new executable
program is generated (compilation and link) for the simulation.

That file name is provided using the keyword: \telkey{FORTRAN FILE}.


\section{The liquid boundaries file}

it is an ASCII file enabling the user to specify time-varying boundary
conditions values (flow rate, depth, velocity, tracer concentration) at all the
liquid boundaries. That file can be generated under the FUDAA-PREPRO software
interface.

For a thorough description of that file, refer to the specific subsection
\ref{sec:liqbnd}.

That name of file is provided using the keyword: \telkey{LIQUID BOUNDARIES
FILE}.


\section{The previous computation file}

it is a \telemac{3D} result file which is used for initializing a new
computation. In order to activate the optional use of that file, the keyword
\telkey{COMPUTATION CONTINUED} should be activated. In order to specify the
previous computation file, its name should be stated through the keyword:
\telkey{PREVIOUS COMPUTATION FILE}. The initial conditions of the new
computation are defined by the last recorded time step in the previous
computation file. The whole set of data from the steering file is read and it
makes possible to redefine or amend the variables (time step, turbulence model,
addition or deletion of a tracer\dots).

A computation can also be initialized from a \telemac{2D} result. In order to
activate that option, the \telkey{2D CONTINUATION} keyword
should be validated. The \telemac{2D} result file should then be associated with
the \telkey{FILE FOR 2D CONTINUATION} keyword (default
value: 'SERAFIN '), see subsection \ref{sec:binfile}.


\section{The reference file}

when a calculation is validated, this file contains the reference result. At
the end of the calculation, the result of the simulation is compared to the
last time step stored in this file. The result of the comparison is given in
the control printout in the form of a maximum difference in elevation and the
three components of velocity.

The name of this file is given by the keyword: \telkey{REFERENCE FILE} and its
format is specified by the keyword: \telkey{REFERENCE FILE FORMAT} (default
value: 'SERAFIN '), see subsection \ref{sec:binfile}.


\section{The stage-discharge curves file}

this text file enables the user to configure the evolution of the prescribed
value on specific open boundaries. This file is used when the prescribed
elevation is determined by a discharge-elevation law. The descriptions of the
appropriate laws are given through this file.

See subsection \ref{sec:discharge} for a complete description of this file.

The name of this file is specified with the keyword: \telkey{STAGE-DISCHARGE
CURVES FILE}.


\section{The sources file}

this text file enables the user to specify values for time-dependent conditions
for sources (discharge, tracers concentration).

See section \ref{sec:srcfile} for a complete description of this file.

The file name is specified with the keyword: \telkey{SOURCES FILE}.


\section{The 3d result file}
\label{sec:3dres}
it is the file into which \telemac{3D} stores the information during the
computation. It is a SERAFIN-formatted file (refer to Appendix
\ref{sec:srffmt}) or a MED-formatted file. It first contains all the data about
mesh geometry, then the names of the stored variables. It also contains, for
each graphic printout and for each mesh point, the values of the various
recorded variables.

Its content varies according to the values of the following keywords:

\telkey{NUMBER OF FIRST TIME STEP FOR GRAPHIC PRINTOUTS}: provided to set from
which time step onwards the information will be stored, in order to prevent too
bulky files, especially when the computation begins with an uninteresting
transient stage related to the definition of unrealistic initial conditions
(e.g. invariably zero currents). The default value is 0 (writing of the graphic
printouts at the beginning of the simulation).

\telkey{GRAPHIC PRINTOUT PERIOD}: sets the period (in number of time steps) of
printouts in order to prevent a too bulky file.  Default value is 1 (writing at
every time step). For instance:

\begin{lstlisting}[language=TelemacCas]
TIME STEP  =  60.0
GRAPHIC PRINTOUT PERIOD =  30
\end{lstlisting}

The results will be backed up every 1,800${}^{th}$ second, i.e. 30${}^{th}$
minute.

\telkey{VARIABLES FOR 3D GRAPHIC PRINTOUTS}: keyword specifying the list of
variables which will be stored in the result file. Each variable is identified
by means of a name from the list below.

\begin{itemize}
\item Z elevation (m),

\item U velocity along \textit{x} axis (m/s),

\item V velocity along \textit{y} axis (m/s),

\item W velocity along \textit{z} axis (m/s),

\item TA concentrations for tracers (TA1 for the 1${}^{st}$ one, TA2 for the
2${}^{nd}$ one\dots ),

\item NUX viscosity for U and V along \textit{x} axis (m${}^{2}$/s),

\item NUY viscosity for U and V along \textit{y} axis (m${}^{2}$/s),

\item NUZ viscosity for U and V along \textit{z} axis (m${}^{2}$/s),

\item NAX viscosity for tracers along \textit{x} axis (m${}^{2}$/s),

\item NAY viscosity for tracers along \textit{y} axis (m${}^{2}$/s),

\item NAZ viscosity for tracers along \textit{z} axis (m${}^{2}$/s),

\item RI Richardson number in case of a mixing length model,

\item K turbulent energy for \textit{k}-$\epsilon$ model (J/kg),

\item EPS dissipation of turbulent energy (W/kg),

\item DP dynamic pressure (multiplied by DT/RHO),

\item PH hydrostatic pressure (Pa),

\item RHO relative density,

\item P1 private variable 1,

\item P2 private variable 2,

\item P3 private variable 3,

\item P4 private variable 4.
\end{itemize}

That file name is provided using the keyword: \telkey{3D RESULT FILE} and its
format using \telkey{3D RESULT FILE FORMAT} (default value: 'SERAFIN '). Stored
variables by default are `Z, U, V, W'.

Using the private arrays requires updating the keyword \telkey{NUMBER OF
PRIVATE ARRAYS} in order to perform the required memory allocations (see
section \ref{sec:privarray}).

\section{The 2d result file}

it is the file into which \telemac{3D} stores the specifically two-dimensional
data during the computation (such as the free surface, the depth-averaged
horizontal components of velocity and the depth-averaged tracers). It has a
SERAFIN format. The free surface and the horizontal components of velocity will
then physically correspond to the same data as those being supplied by
\telemac{2D}. The obtained values, however, may be different from an analogue
computation being directly made with \telemac{2D} when the flow is specifically
three-dimensional.

Its content varies according to the values of the following keywords:

\telkey{NUMBER OF FIRST TIME STEP FOR GRAPHIC PRINTOUTS}: same keyword as that
described in section \ref{sec:3dres}.

\telkey{GRAPHIC PRINTOUT PERIOD}: same keyword as that described in section
3.11.

\telkey{VARIABLES FOR 2D GRAPHIC PRINTOUTS}: keyword specifying the list of
variables which will be stored in the result file. Each variable is identified
by means of a name from the list below.

\begin{itemize}
\item U average velocity along \textit{x} axis (m/s),

\item V average velocity along \textit{y} axis (m/s),

\item C celerity (m/s),

\item H water depth (m),

\item S free surface elevation (m),

\item B bottom elevation (m),

\item TA averaged concentrations for tracers (TA1 for the 1${}^{st}$ one, TA2
for the 2${}^{nd}$ one\dots ),

\item F Froude number (m),

\item Q scalar discharge (m${}^{2}$/s),

\item I discharge along \textit{x} (m${}^{2}$/s),

\item J discharge along \textit{y} (m${}^{2}$/s),

\item M norm of velocity (m/s),

\item X wind along \textit{x} axis (m/s),

\item Y wind along \textit{y} axis (m/s),

\item P air pressure (Pa),

\item W friction coefficient,

\item RB non erodible bottom elevation (m),

\item HD thickness of the sediment bed layer (m),

\item EF erosion rate (kg/m${}^{3}$/s),

\item DF deposition flux (kg/m${}^{3}$/s),

\item DZF bed evolution,

\item PRIVE1 private array PRIVE 1,

\item PRIVE2 private array PRIVE 2,

\item PRIVE3 private array PRIVE 3,

\item PRIVE4 private array PRIVE 4,

\item QS  solid discharge (m${}^{2}$/s),

\item QSX  solid discharge along \textit{x} (m${}^{2}$/s),

\item QSY  solid discharge along \textit{y} (m${}^{2}$/s),

\item US  friction velocity (m/s),

\item MAXZ  maximum value of the water elevation during the computation (m),

\item TMXZ  time corresponding to this maximum elevation (s).
\end{itemize}

That file name is provided by means of the keyword: \telkey{2D RESULT FILE} and
its format using \telkey{2D RESULT FILE FORMAT} (default value: `SERAFIN ').
Stored variables by default are `U, V, H, B'.

Using the private arrays requires updating the keyword \telkey{NUMBER OF
PRIVATE ARRAYS} in order to perform the required memory allocations (see
section \ref{sec:privarray}).

\section{The output listing}

it is a formatted file which can be created by \telemac{3D} during the
computation (program launched with the -s option). It contains the report of a
\telemac{3D} running. Its contents vary according to the values of the following
keywords:

\telkey{NUMBER OF FIRST TIME STEP FOR LISTING PRINTOUTS}: provided to set at
which time step it is desired to begin printing the data, in order to prevent
too bulky files.

\telkey{LISTING PRINTOUT PERIOD}: sets the period between two time step
printings. The value is given as a time step number. For instance, the
following sequence:

\begin{lstlisting}[language=TelemacCas]
TIME STEP = 30.0
LISTING PRINTOUT PERIOD = 2
\end{lstlisting}

Prints the output listing every minute of simulation.

\telkey{MASS-BALANCE}: if it is requested, the user will get
information about the mass flow (or rather the volumes) and the errors
(primarily linked to the precision achieved by the solvers) of that computation
in the domain. This is not done by default.

\telkey{INFORMATION ABOUT MASS-BALANCE FOR EACH LISTING PRINTOUT}: if they are
requested (that is done by default), the user will get, at each time step,
information about the flows within the domain.

The file name is directly managed by the \telemac{3D} launching procedure.
Generally, the file has a name which is created from the name in the steering
file and the number of the processors used in the computation, followed by the
suffix ".sortie".

\section{The file for scope}

tELEMAC-3D gives the user an opportunity to retrieve computation variables
along a 1D profile during a computation. These profiles will then be stored in
a dedicated file the name of which is defined by the keyword \telkey{FILE FOR
SCOPE}. For that purpose, the user should program the SCOPE subroutine.

That subroutine is provided to define profiles of computed variables, or other
variables, as created by a user along a segment with (X1, Y1, Z1) and (X2, Y2,
Z2) co-ordinates. The user also decides the number of points as distributed
along that segment. The data is automatically saved as per the SCOPE format at
all the time steps (for a description of this format, see Appendix of RUBENS
Version 4.0. Manuel Utilisateur, O. Quiquempoix, EDF report HE-45/95/031/B).

\section{The auxiliary files}

Other files can be used by \telemac{3D}:

\begin{itemize}
\item One or two binary data files, as specified by the keywords
\telkey{BINARY DATA FILE 1} and \telkey{BINARY DATA FILE 2}. These files can be
used to provide data to the program, the user has to handle their reading
within the \telkey{FORTRAN FILE}. The data from these files shall be read using
the T3DBI1 logic unit for binary data file 1 and the T3DBI2 logic unit for
binary data file 2,

\item One or two formatted data files, as specified by the keywords
\telkey{FORMATTED DATA FILE 1} and \telkey{FORMATTED DATA FILE 2}. These files
can be used to provide data to the program, the user has to handle their
reading within the \telkey{FORTRAN FILE}. The data from these files shall be
read using the T3DFO1 logic unit for formatted data file 1 and the T3DFO2 logic
unit for formatted data file 2,

\item A binary results file, as specified by the keyword \telkey{BINARY
RESULTS FILE}. \telkey{BINARY RESULTS FILE}This file can be used to store
additional results, the user has to handle their writing within the Fortran
program using the T3DRBI logic unit.

\item A formatted results file, as specified by the keyword \telkey{FORMATTED
RESULTS FILE}. This file can be used to store additional results (for instance,
results readable by an external 1D simulation code used for a coupling with
TELEMAC), the user has to handle their writing within the Fortran program using
the T3DRBO logic unit.
\end{itemize}

The read or write operations from/into these files should be thoroughly managed
by the user (the files are opened and closed by the program). That management
can be performed from any point which the user can gain access to.  The logic
unit numbers are stated in the \telfile{DECLARATIONS\_TELEMAC3D} module and the user can
access them through a \telfile{USE DECLARATIONS\_TELEMAC3D} command at the beginning of a
subroutine. For instance, using a file for providing the initial conditions
will result in managing that file within the CONDIM subroutine. Likewise, using
a file for inserting boundary conditions will be possible at the BORD3D
subroutine. In case of conflicting statements, one can use, for example:
\telfile{USE DECLARATIONS\_TELEMAC3D, ONLY: T3DBI1}.

\section{The dictionnary file}

This is a file containing all the information on the keywords (French name,
English name, default value, type, keywords documentation). This file can be
read by the user, but it should not be modified in any case.

\section{Topographic and bathymetric data}
\label{sec:topo}
Topographical and bathymetric data may be supplied to \telemac{3D} at three
levels:

\begin{itemize}
\item Either directly in the geometry file by a topographical or bathymetric
value associated with each mesh node. In this case, the data are processed
while the mesh is being built using MATISSE, JANET or BLUE KENUE,MATISSE or
when the \bief module is run before \telemac{3D} is started. \bief reads
the information in one or more bottom topography files (5 at most) and
interpolates at each point in the domain,

\item Or in the form of a cluster of points with elevations that have no
relation with the mesh nodes, during the \telemac{3D} computation. \telemac{3D}
then makes the interpolation directly with the same algorithm as \bief. The
file name is provided by the keyword \telkey{BOTTOM TOPOGRAPHY FILE}. Unlike
\bief, \telemac{3D} only manages one single bottom topography file. This may be
in SINUSX format or more simply a file consisting of three columns X, Y, Z. The
SINUSX format is described in the RUBENS user manual,

\item Or using the \telfile{T3D\_CORFON} subroutine. This is usually used for schematic
test cases.
\end{itemize}

In all cases, \telemac{3D} offers the possibility of smoothing the bottom
topography in order to obtain a more regular geometry. The smoothing algorithm
can be iterated several times depending on the degree of smoothing required.
The keyword \telkey{NUMBER OF BOTTOM SMOOTHINGS} then defines the number of
iterations carried out in the \telfile{T3D\_CORFON} subroutine. The default
value of this keyword is 0.
