% \section{Bibliography}

\pagebreak
\bibliographystyle{plainnat}

\begin{thebibliography}{99}

%biblio myModule
\bibitem{estelprinciple}  {\sc Decung F., Falappi S., Vitorge E.},\\
 {\em \estel version \rel: principle note},\\
H-P78-2012-xxx, EDF R\&D, LNHE, 2012.

\bibitem{esteluser}  {\sc Decung F., Falappi S., Vitorge E.},\\
 {\em \estel version \rel: user manual},\\
H-P78-2012-xxx, EDF R\&D, LNHE, 2012.

\bibitem{estelvalidation}  {\sc Decung F., Falappi S., Vitorge E.},\\
 {\em \estel version \rel: validation note},\\
H-P78-2012-xxx, EDF R\&D, LNHE, 2012.

\bibitem{estelqualification}  {\sc Falappi S., Sarotte L.},\\
 {\em \estel version \rel: qualification note},\\
H-P78-2012-xxx, EDF R\&D, LNHE, 2012.

%biblio openTELEMAC
\vspace{0.5cm}
{\bf Biblio openTELEMAC}

\bibitem{JMH1}  {\sc Hervouet J.-M., P. Bates},\\
 {\em TELEMAC modelling system: an overview},\\
Hydrological Processes, Vol. 14, 2000.
% 
\bibitem{JMH2}  {\sc Hervouet J.-M.},\\
 {\em Hydrodynamics of free surface flows: modelling with finite element methods},\\
John Wiley \& Sons, 2007.
% 
\bibitem{bief}  {\sc Hervouet J.-M.},\\
 {\em Guide to programming in the TELEMAC system - Version 6.0},\\
H-P74-2009-00801-EN, EDF R\&D, LNHE, 2009.

%biblio Couplage
\vspace{0.5cm} {\bf Biblio Coupling}
% 
\bibitem{renaud-HIC}  {\sc Renaud J-P., Cloke A. and Anderson M.},\\
 {\em A new coupled model for sudying the hyporheic zone and hillslope-floodplain interactions},\\
7th International Conference on Hydroinformatics, HIC 2006, Nice, France.

\bibitem{claxton}  {\sc Claxton A., Bates P. and Cloke A.},\\
 {\em Mixing of hillslope, river, and alluvial ground waters in lowland floodplains},\\
Ground Water, Volume 41, Issue 7, p.926-936, 2003.

% 
\end{thebibliography}
