% \section{Bibliography}

\pagebreak
\bibliographystyle{plainnat}

\begin{thebibliography}{99}

%biblio myModule
%\bibitem{test}  {\sc author},\\
% {\em title},\\
%Other stuff, Year.

\bibitem{alves} {\sc ALVES J.H.G.M., BANNER M.L.},\\ 
{\em  Revisiting the Pierson-Moskowitz Asymptotic Limits for Fully Developed Wind Waves,}\\
Journal of Physical Oceanography, volume 33, pp 1301 - 1323, 2003.

\bibitem{battjes}{\sc  BATTJES J.A., JANSSEN J.P.F.M.},\\ 
{\em Energy loss and set-up due to breaking of random waves. }\\
Proc.16th Int. Conf. on Coastal Eng., pp 569-587, 1978.

\bibitem{benoit}{\sc   M. BENOIT},\\ 
{\em  Fiche cas-test : G\'en\'eration d'\'etats de mer en fonction de la distance d'action du vent (fetch) }\\
Extrait de la note HE-42/96/010/B. Logiciel TOMAWAC de mod\'elisation des \'etats de mer en \'el\'ements finis, Dossier de validation de la version 1.0, 1996.  


\bibitem{CERC77}{\sc Coastal Engineering Research Center (CERC)},\\  
{\em Shore Protection Manual. }
USACE, Vicksburg, MS, 1977.\\

\bibitem{CERC84}{\sc Coastal Engineering Research Center (CERC)},\\  
{\em Shore Protection Manual Volume 1. }\\
USACE, Vicksburg, MS, 1984.


\bibitem{gagnaire} {\sc GAGNAIRE-RENOU E., BENOIT M., FORGET PH.},\\  
{\em Ocean wave spectrum properties as derived from quasi-exact computations of nonlinear wave-wave interactions. }\\
J. Geophys. Res. C (Oceans), Vol. 115, C12, C12058, doi:10.1029/2009JC005665, 2010

\bibitem{haselmann1} {\sc K Hasselmann, TP Barnett, E Bouws, H Carlson},\\  
{\em  Measurements of wind-wave growth and swell decay during the Joint North Sea Wave Project (JONSWAP),} \\
 Dtsh. Hydrogh. Z., Suppl. A, 8, 12, 95 pp, 1973.

\bibitem{haselmann2} {\sc HASSELMANN S., HASSELMANN K.},\\  
{\em Computations and parameterizations of the nonlinear energy
transfer in gravity-wave spectrum. Part I: a new method for efficient computations of the exact nonlinear transfer integral,}\\
 J. Phys. Oceanogr., vol. 15, pp 1369-1377,1985.

\bibitem{haselmann3} {\sc HASSELMANN S., HASSELMANN K., ALLENDER J.H., BARNETT T.P..},\\  
{\em Computations and parameterizations of the nonlinear energy transfer in gravity-wave spectrum. Part II: Parameterizations of the nonlinear energy transfer for application in wave models.} \\
J. Phys. Oceanogr., vol. 15, pp 1378-1391, 1985.

\bibitem{janssen89} {\sc JANSSEN P.A.E.M.},\\  
{\em   (1989): Wave-induced stress and the drag of air flow over sea waves.}\\
 J. Phys. Oceanogr., vol. 19, pp 745-754, 1989.

\bibitem{janssen91} {\sc JANSSEN P.A.E.M.},\\  
{\em   (1991): Quasi-linear theory of wind-wave generation applied to wave forecasting. }\\
J. Phys. Oceanogr., vol. 21, pp 1631-1642, 1991.

\bibitem{kahma} {\sc   K.K. KAHMA \& C J. CALKOEN},\\  
{\em Reconciling discrepancies in the observed growth of wind-generated waves.}\\
 J. Phys. Oceanogr., vol. 22, pp. 1389-1405,1992.

\bibitem{komen1} {\sc  KOMEN G.J., HASSELMANN S., HASSELMANN K.}\\
{\em On the existence of a fully developed wind-sea spectrum. }\\
J. Phys. Oceanogr., vol. 14, pp 1271-1285, 1984.

\bibitem{komen2} {\sc  KOMEN G.J., CAVALERI L., DONELAN M., HASSELMANN K., HASSELMANN S., JANSSEN P.A.E.M.}\\
{\em Dynamics and Modelling of ocean waves.}\\
 Cambridge University Press, 1994,532 p.

\bibitem{snyder} {\sc  SNYDER. R.L., DOBSON F.W., ELLIOT J.A., LONG R.B.}\\
{\em Array measurements of atmospheric pressure fluctuations above surface gravity waves.}\\ 
 J. Fluid Mech., vol. 102, pp 1-59, 1981.

\bibitem{WESTHUYSEN_2007} {\sc  Van der WESTHUYSEN A. J., ZIJLEMA M., BATTJES J.A.}\\
{\em Nonlinear saturation-based whitecapping dissipation in swan for deep and shallow water.}\\
 Coastal Eng., 54, 151-170, 2007.

\bibitem{wilson65} {\sc B W. WILSON}\\
{\em Numerical prediction of ocean waves in the North Atlantic for December 1959.}\\
 Deutsche Hydrographische Zeitschrift, vol. 18 (3), pp. 114-130, 1965.

\bibitem{Yan} {\sc YAN L. }\\
{\em An improved wind input source term for third generation ocean wave modelling.}\\
 Tech. Rep. 8, Royal Dutch Meteor. Inst, 1987.

\bibitem{young} {\sc  I R YOUNG \& L A. VERHAGEN}\\
{\em The growth of fetch-limited waves in water of finite depth. Part I: total energy and peak frequency.}\\
 Coastal Engg. vol. 28, pp. 47-78, 1996.

\bibitem{vincent} {\sc   VINCENT C.L.,BRIGGS M.J.}\\
{\em Refraction-diffraction of irregular waves over a mound. }\\
Journal os Waterway, Port,Coastal and Ocean Engineering, Vol.115, N2, pp 269-284, 1989

\bibitem{mattarolo} {\sc   MATTAROLO G.}\\
{\em  Simulating diffraction in the third generation spectral model TOMAWAC. Implementation and testing of a model based on previous research work. }\\
H-P74-2012-00439-EN, pp19-48, 2012

\bibitem{becq} {\sc    F. BECQ, M. BENOIT }\\
{\em Fiche cas-test : Propagation de houle sur un haut fond elliptique.}\\
Extrait de la note HE-42/96/010/B Logiciel TOMAWAC de modelisation des tats de mer en elements finis, Dossier de validation de la version 1.0, 1996. 

\bibitem{hedges}  {\sc HEDJES TS, ANASTASIOU K, GABRIEL D},\\ 
{\em  Interaction of random waves and currents},\\
 J Waterway, Port, Coastal, Ocean Eng., 111(2), 275–288, 1985.

\bibitem{lai}  {\sc LAI RJ, LONG SR, HUANG N.E.},\\
 {\em  Laboratory studies of wave-current interaction : kinematics of the strong interaction.},\\
 Journal of Geophysical Research 94, p 16201-16214, 1989.

\bibitem{Westhuysen_2012}  {\sc Van der WESTHUYSEN A. J.},\\
 {\em Spectral modeling of wave dissipation on negative current gradients.},\\
Coastal Eng, 58 P 17-30, 2012.

\bibitem{mathiesen} {\sc MATHIESEN M. },\\
 {\em Wave refraction by a current whirl.}\\
 Journal of Geophysical Research, vol. 92, n C4, pp 3905-3912, april 15, 1987.

\bibitem{tholman} {\sc TOLMAN H.L. },\\
 {\em  A third-generation model for wind waves on slowly varying, unsteady and inhomogeneous depths and currents.}\\
 Journal of Physical Oceanography, Vol. 21, n 6, june 1990.

\bibitem{hubbert} {\sc HUBBERT K.P. et. Al. },\\
 {\em Numerical investigation of depth and current refraction of waves,}\\
 Journal of Geophysical Research, vol. 96, n C2, pp 2737-2748, february 15, 1991.

\bibitem{cowadis} {\sc MarcosF. Benoit M., Thellier P.},\\
 {\em Cowadis software for finite element wave propagation in coastal areas : Validation document of release 1.0}\\
Tecnical report EDF R\&D HE-42/98/057/A, 1998.

\bibitem{suzuki} {\sc T. Suzuki, M. Zijlema, B. Burger, M. C. Meijer, S. Narayan},\\
 {\em Wave dissipation by vegetation with layer schematisation in SWAN}\\
Coast. Eng., vol. 59, pp. 64-71, 2011.

\bibitem{mendez}{\sc  F. M. Mendez and I. J. Losada, },\\
 {\em An empirical model to estimate the propagation of random breaking and nonbreaking waves over vegetation fields}\\
 Coast. Eng., vol. 51, pp. 103-118, 2004.

\bibitem{swamp85}{\sc Sea Wave Modelling Project (SWAMP) Groupe} ,\\
 {\em An Intercomparison study of wind wave prediction models. Part 1 : Principal results and conclusions}\\
Ocean Wave Modeling, Plenum Press, New York, 256 p, 1985.  

\bibitem{vledder} {\sc VAN VLEDDER G. Ph.},\\
 {\em Directional response of wind waves to turning winds. }\\
Thèse de Doctorat. Rapport 90-2. Delft University of Technology (Pays-Bas), 1990.

\bibitem{valid10TW}{\sc   M. BENOIT},\\ 
{\em  Fiche cas-test :Evolution du spectre directionnel de houle dans un vent tournant }\\
Extrait de la note HE-42/96/010/B. Logiciel TOMAWAC de mod\'elisation des \'etats de mer en \'el\'ements finis, Dossier de validation de la version 1.0, 1996.  

%biblio openTELEMAC
\vspace{0.5cm}
{\bf Biblio openTELEMAC}

\bibitem{JMH1}  {\sc Hervouet J.-M., P. Bates},\\
 {\em TELEMAC modelling system: an overview},\\
Hydrological Processes, Vol. 14, 2000.
% 
\bibitem{JMH2}  {\sc Hervouet J.-M.},\\
 {\em Hydrodynamics of free surface flows: modelling with finite element methods},\\
John Wiley \& Sons, 2007.
% 
\bibitem{bief}  {\sc Hervouet J.-M.},\\
 {\em Guide to programming in the TELEMAC system - Version 6.0},\\
H-P74-2009-00801-EN, EDF R\&D, LNHE, 2009.

%biblio Couplage
\vspace{0.5cm} {\bf Biblio Coupling}
% 
\bibitem{renaud-HIC}  {\sc Renaud J-P., Cloke A. and Anderson M.},\\
 {\em A new coupled model for sudying the hyporheic zone and hillslope-floodplain interactions},\\
7th International Conference on Hydroinformatics, HIC 2006, Nice, France.

\bibitem{claxton}  {\sc Claxton A., Bates P. and Cloke A.},\\
 {\em Mixing of hillslope, river, and alluvial ground waters in lowland floodplains},\\
Ground Water, Volume 41, Issue 7, p.926-936, 2003.

\bibitem{vito}  {\sc V. Bacchi  E. Gagnaire-Renou N. Durand M. Benoit},\\
 {\em Wave energy dissipation by vegetation in TOMAWAC},\\
21st Telemac- Mascaret User Club Grenoble, France, 15-17 October, 2014

\bibitem{dean91} {\sc R.G Dean},\\
 {\em Equilibrium beach profiles: characteristics and applications},\\
Journal of Coastal Research, 7 (1), pp. 53–84, 1991
% 
\end{thebibliography}
